---
title: "Priming Free Recall"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading the Data

```{r}
initialrecall = read.csv("sem_compiled_initial_2020.csv",header = TRUE, sep = ",")
library(tidyverse)

mean_duration = group_by(initialrecall, participant) %>%
  summarise_at(vars(PrimeDuration), mean, na.rm = TRUE)

initialrecall$Studied[is.na(initialrecall$Studied)] = 0
initialrecall$Acc = ifelse(initialrecall$Studied == "Studied", 1,0)
```

## Means Per Condition

```{r}
mean_subject = group_by(initialrecall, participant, PrimeCondition) %>%
  summarise_at(vars(Acc), mean)
mean_cond = Rmisc::summarySE(mean_subject,
                             measurevar = "Acc",
                             groupvars = "PrimeCondition", 
                             na.rm = TRUE)

mean_cond = mean_cond %>% filter(PrimeCondition != "firstword")

library(ggplot2)
library(ggthemes)
mean_cond %>% 
  ggplot(aes(x = PrimeCondition, y = Acc))+
geom_bar(stat = "identity", position = "dodge", width = 0.7, fill = "darkgoldenrod",
         color = "gray26")+
  geom_errorbar(aes(ymin=Acc - se, ymax=Acc + se), 
             width=.2, color = "gray26", 
             position = position_dodge(0.7))+
 theme_few()+
  scale_fill_gdocs()+
    xlab("Prime Condition") + ylab("Mean Accuracy (Correctly Recalled Items)") + 
  ggtitle("Free Recall Online: Immediate Test")  +
   theme(axis.text = element_text(size = rel(1)),
          axis.title = element_text(face = "bold", size = rel(1)),
          legend.title = element_text(face = "bold", size = rel(1)),
         plot.title = element_text(hjust = .5),
         strip.text.x = element_text(face = "bold", size = rel(1.4)))

## anova/t-test code

mean_cond_sub = group_by(initialrecall, participant, PrimeCondition) %>%
  summarize_at(vars(Acc), mean)

mean_cond_sub_main = mean_cond_sub %>% filter(PrimeCondition != "firstword")

## anova code
mean_cond_sub_main$participant = as.factor(mean_cond_sub_main$participant)
aov_prime = aov(data = mean_cond_sub_main, 
                Acc ~ PrimeCondition + Error(participant/PrimeCondition))
summary(aov_prime)

## p and F value will tell you if there is 
## a main effect of prime condition

## t-test

ident_sub = mean_cond_sub %>% filter(PrimeCondition == "Identical")
related_sub = mean_cond_sub %>% filter(PrimeCondition == "Related")
unrelated_sub = mean_cond_sub %>% filter(PrimeCondition == "Unrelated")

# ident vs. unrelated
t.test(ident_sub$Acc, unrelated_sub$Acc, paired = TRUE)
# related vs. unrelated
t.test(related_sub$Acc, unrelated_sub$Acc, paired = TRUE)
# ident vs. related
t.test(ident_sub$Acc, related_sub$Acc, paired = TRUE)

```

# Delayed Recognition Trials

```{r}
delayedrecog = read.csv("sem_compiled_recog_2020.csv", header = TRUE, sep = ",")

delayedrecog$WordStatus = ifelse(delayedrecog$WordStatus == "Old", "Word was: Old", 
                                 "Word was: New")

recog_acc_subject = group_by(delayedrecog, participant, WordStatus, Day1Recalled, RecogDecision) %>%
  summarise(Trials = n())

recog_rmisc = Rmisc::summarySE(recog_acc_subject,
                               measurevar = "Trials",
                               groupvars = c("WordStatus", "Day1Recalled", "RecogDecision"))

recog_rmisc %>% 
  ggplot(aes(x = RecogDecision, y = Trials, 
             group =Day1Recalled , fill = Day1Recalled))+
geom_bar(stat = "identity", position = "dodge", width = 0.7)+
  geom_errorbar(aes(ymin=Trials - se, ymax=Trials + se), 
             width=.2, color = "gray26", 
             position = position_dodge(0.7))+
 theme_few()+
  facet_wrap(~WordStatus)+
  scale_fill_gdocs()+
    xlab("Recognition Decision") + ylab("Mean Number of Trials") + 
  ggtitle("Delayed Recognition Test")  +
   theme(axis.text = element_text(size = rel(1)),
          axis.title = element_text(face = "bold", size = rel(1)),
          legend.title = element_text(face = "bold", size = rel(1)),
         plot.title = element_text(hjust = .5),
         strip.text.x = element_text(face = "bold", size = rel(1.4)))
```

# Hits, FA, CR, and Miss

## Partitioning the Data

```{r}

delayedrecog$WithinTrialPrimeCondition = ifelse(is.na(delayedrecog$WithinTrialPrimeCondition), "Unprimed", delayedrecog$WithinTrialPrimeCondition)

delayedrecog$CrossTrialPrimeCondition = ifelse(is.na(delayedrecog$CrossTrialPrimeCondition), "Unprimed", delayedrecog$CrossTrialPrimeCondition)

notstudied = delayedrecog %>% filter(WordStatus == "Word was: New")
## now these "new" items could have been within/cross-trial primed
## although "within" is probably very few trials

## we look at false alarms and correct rejections of these "new" items

studied = delayedrecog %>% filter(WordStatus == "Word was: Old")
## now these "old" items could have been within/cross-trial primed

## we look at hits and misses of these "OLD" items

```

## Old (Studied) Items

### Within-Trial priming
```{r}
studied_withintrial_measures = group_by(studied, participant, WithinTrialPrimeCondition) %>% 
  summarise_at(vars(Hit, Miss), mean)

studied_withintrial_measures_long = gather(studied_withintrial_measures, 
                                   Measure, Proportion, Hit, Miss)

studied_withintrial_rmisc = Rmisc::summarySE(studied_withintrial_measures_long,
                                     measurevar = "Proportion",
                        groupvars = c( "WithinTrialPrimeCondition",
                                                   "Measure"))

studied_withintrial_rmisc = studied_withintrial_rmisc %>%
  filter(!WithinTrialPrimeCondition %in% "firstword")

studied_withintrial_rmisc %>% 
  ggplot(aes(x = Measure, y = Proportion, 
             group =WithinTrialPrimeCondition , fill = WithinTrialPrimeCondition))+
  #geom_point()+
geom_bar(stat = "identity", position = "dodge", width = 0.7)+
  geom_errorbar(aes(ymin=Proportion - se, ymax=Proportion + se), 
             width=.2, color = "gray26", 
             position = position_dodge(0.7))+
 theme_few()+
  scale_fill_gdocs()+
    xlab("") + ylab("Proportion") + 
  ggtitle("OLD items: within-trial priming")  +
   theme(axis.text = element_text(size = rel(1)),
          axis.title = element_text(face = "bold", size = rel(1)),
          legend.title = element_text(face = "bold", size = rel(1)),
         plot.title = element_text(hjust = .5),
         strip.text.x = element_text(face = "bold", size = rel(1.4)))

## checking if hit/miss rate is reliably different between ident and rel/unrel

hit_data = studied_withintrial_measures_long %>% filter(Measure == "Hit" & 
                              !WithinTrialPrimeCondition %in% c("Unprimed", "firstword"))

hit_aov = aov(data = hit_data, Proportion ~ WithinTrialPrimeCondition + 
                Error(participant/WithinTrialPrimeCondition))
summary(hit_aov)

hit_ident = hit_data %>% filter(WithinTrialPrimeCondition == "Identical")
hit_rel = hit_data %>% filter(WithinTrialPrimeCondition == "Related")
hit_unrel = hit_data %>% filter(WithinTrialPrimeCondition == "Unrelated")

t.test(hit_ident$Proportion, hit_rel$Proportion, paired = TRUE)
t.test(hit_ident$Proportion, hit_unrel$Proportion, paired = TRUE)

miss_data = studied_withintrial_measures_long %>% filter(Measure == "Miss" & 
                              !WithinTrialPrimeCondition %in% c("Unprimed", "firstword"))

miss_aov = aov(data = miss_data, Proportion ~ WithinTrialPrimeCondition + 
                Error(participant/WithinTrialPrimeCondition))
summary(miss_aov)

miss_ident = miss_data %>% filter(WithinTrialPrimeCondition == "Identical")
miss_rel = miss_data %>% filter(WithinTrialPrimeCondition == "Related")
miss_unrel = miss_data %>% filter(WithinTrialPrimeCondition == "Unrelated")

t.test(miss_ident$Proportion, miss_rel$Proportion, paired = TRUE)
t.test(miss_ident$Proportion, miss_unrel$Proportion, paired = TRUE)

```
## HIT - MISS sensitivity
```{r}
studied_withintrial_measures$HitMinusMiss = studied_withintrial_measures$Hit - 
  studied_withintrial_measures$Miss

hitminusmiss_maindata = studied_withintrial_measures %>% 
  filter(!WithinTrialPrimeCondition %in% c("firstword", "Unprimed"))

hitminusmiss_maindata_aov = aov(data = hitminusmiss_maindata,
                                HitMinusMiss ~ WithinTrialPrimeCondition + 
                                  Error(participant/WithinTrialPrimeCondition))

summary(hitminusmiss_maindata_aov)


hitminusmiss_maindata_ident = hitminusmiss_maindata %>% 
  filter(WithinTrialPrimeCondition == "Identical")
hitminusmiss_maindata_rel = hitminusmiss_maindata %>% 
  filter(WithinTrialPrimeCondition == "Related")
hitminusmiss_maindata_unrel = hitminusmiss_maindata %>% 
  filter(WithinTrialPrimeCondition == "Unrelated")

t.test(hitminusmiss_maindata_ident$HitMinusMiss, hitminusmiss_maindata_unrel$HitMinusMiss,
       paired = TRUE)

t.test(hitminusmiss_maindata_ident$HitMinusMiss, hitminusmiss_maindata_rel$HitMinusMiss,
       paired = TRUE)


hitminusmiss_rmisc = Rmisc::summarySE(hitminusmiss_maindata,
                                     measurevar = "HitMinusMiss",
                        groupvars = c( "WithinTrialPrimeCondition"))


hitminusmiss_rmisc %>% 
  ggplot(aes(x = WithinTrialPrimeCondition, y = HitMinusMiss, 
             group =WithinTrialPrimeCondition , fill = WithinTrialPrimeCondition))+
  #geom_point()+
geom_bar(stat = "identity", position = "dodge", width = 0.7)+
  geom_errorbar(aes(ymin=HitMinusMiss - se, ymax=HitMinusMiss + se), 
             width=.2, color = "gray26", 
             position = position_dodge(0.7))+
 theme_few()+
  scale_fill_gdocs()+
    xlab("") + ylab("Difference Proportion") + 
  ggtitle("OLD items: within-trial priming \n Hits - Misses")  +
   theme(axis.text = element_text(size = rel(1)),
          axis.title = element_text(face = "bold", size = rel(1)),
          legend.title = element_text(face = "bold", size = rel(1)),
         plot.title = element_text(hjust = .5),
         strip.text.x = element_text(face = "bold", size = rel(1.4)))
```
### Cross-Trial priming
```{r}
studied_crosstrial_measures = group_by(studied, participant, CrossTrialPrimeCondition) %>%
  summarise_at(vars(Hit, Miss), mean)

studied_crosstrial_measures_long = gather(studied_crosstrial_measures, 
                                   Measure, Proportion, Hit, Miss)

studied_crosstrial_rmisc = Rmisc::summarySE(studied_crosstrial_measures_long,
                                     measurevar = "Proportion",
                        groupvars = c("CrossTrialPrimeCondition",
                                                   "Measure"))

studied_crosstrial_rmisc %>% 
  ggplot(aes(x = Measure, y = Proportion, 
             group =CrossTrialPrimeCondition , fill = CrossTrialPrimeCondition))+
  #geom_point()+
geom_bar(stat = "identity", position = "dodge", width = 0.7)+
  geom_errorbar(aes(ymin=Proportion - se, ymax=Proportion + se), 
             width=.2, color = "gray26", 
             position = position_dodge(0.7))+
 theme_few()+
  scale_fill_gdocs()+
    xlab("") + ylab("Proportion") + 
  ggtitle("OLD items: cross-trial priming")  +
   theme(axis.text = element_text(size = rel(1)),
          axis.title = element_text(face = "bold", size = rel(1)),
          legend.title = element_text(face = "bold", size = rel(1)),
         plot.title = element_text(hjust = .5),
         strip.text.x = element_text(face = "bold", size = rel(1.4)))
```

# New (Not Studied) Items

### Within-Trial priming
```{r}
withintrial_measures = group_by(notstudied, participant, WithinTrialPrimeCondition) %>% 
  summarise_at(vars(FalseAlarm, CorrectRejection), mean)

withintrial_measures_long = gather(withintrial_measures, 
                                   Measure, Proportion, FalseAlarm, CorrectRejection)

withintrial_rmisc = Rmisc::summarySE(withintrial_measures_long,
                                     measurevar = "Proportion",
                                     groupvars = c("WithinTrialPrimeCondition",
                                                   "Measure"))

withintrial_rmisc %>% 
  ggplot(aes(x = Measure, y = Proportion, 
             group =WithinTrialPrimeCondition , fill = WithinTrialPrimeCondition))+
  #geom_point()+
geom_bar(stat = "identity", position = "dodge", width = 0.7)+
  geom_errorbar(aes(ymin=Proportion - se, ymax=Proportion + se), 
             width=.2, color = "gray26", 
             position = position_dodge(0.7))+
 theme_few()+
  scale_fill_gdocs()+
    xlab("") + ylab("Proportion") + 
  ggtitle("NEW items: within-trial priming")  +
   theme(axis.text = element_text(size = rel(1)),
          axis.title = element_text(face = "bold", size = rel(1)),
          legend.title = element_text(face = "bold", size = rel(1)),
         plot.title = element_text(hjust = .5),
         strip.text.x = element_text(face = "bold", size = rel(1.4)))
```
### Cross-Trial priming
```{r}
crosstrial_measures = group_by(notstudied, participant, CrossTrialPrimeCondition) %>%
  summarise_at(vars(FalseAlarm, CorrectRejection), mean)

crosstrial_measures_long = gather(crosstrial_measures, 
                                   Measure, Proportion, FalseAlarm, CorrectRejection)

crosstrial_rmisc = Rmisc::summarySE(crosstrial_measures_long,
                                     measurevar = "Proportion",
                        groupvars = c("CrossTrialPrimeCondition",
                                                   "Measure"))

crosstrial_rmisc %>% 
  ggplot(aes(x = Measure, y = Proportion, 
             group =CrossTrialPrimeCondition , fill = CrossTrialPrimeCondition))+
  #geom_point()+
geom_bar(stat = "identity", position = "dodge", width = 0.7)+
  geom_errorbar(aes(ymin=Proportion - se, ymax=Proportion + se), 
             width=.2, color = "gray26", 
             position = position_dodge(0.7))+
 theme_few()+
  scale_fill_gdocs()+
    xlab("") + ylab("Proportion") + 
  ggtitle("NEW items: cross-trial priming")  +
   theme(axis.text = element_text(size = rel(1)),
          axis.title = element_text(face = "bold", size = rel(1)),
          legend.title = element_text(face = "bold", size = rel(1)),
         plot.title = element_text(hjust = .5),
         strip.text.x = element_text(face = "bold", size = rel(1.4)))
```



# Recognition RTs

```{r}
## convert RTs to ms
delayedrecog$RT = delayedrecog$key_resp_3.rt*1000
```
## z-score RTs
```{r}
## z-score RTs at participant level
## trim RTs
sem_firsttrim = delayedrecog %>% filter( RT > 50 & RT < 2500)
## aggregate per subject 
meanRT = group_by(sem_firsttrim, participant) %>%
  summarise_at(vars(RT), mean)
colnames(meanRT) = c("participant", "MeanRT")

sdRT = group_by(sem_firsttrim, participant) %>%
  summarise_at(vars(RT), sd)
colnames(sdRT) = c("participant", "sdRT")

RT_agg = merge(meanRT, sdRT, by = "participant")

## merge aggregate info with long data
sem_z = merge(sem_firsttrim, RT_agg, by = "participant", all.x = T)

## person and grand-mean centered scores using original and aggregate
library(dplyr)
sem_z = sem_z %>% mutate(zRT = (RT - MeanRT)/sdRT)
                 
## checking: subject level means should be zero

sub_pic = group_by(sem_z, participant) %>%
  summarise_at(vars(zRT), mean)

## trimming above and below 3 sd
sem_z_trimmed = subset(sem_z, sem_z$zRT < 3 & sem_z$zRT > -3)

## aggregate per subject all IVs and DVs
meanRT_trim = group_by(sem_z_trimmed, participant) %>%
  summarise_at(vars(RT), mean)
colnames(meanRT_trim) = c("participant", "MeanRT_trim")

sdRT_trim = group_by(sem_z_trimmed, participant) %>%
  summarise_at(vars(RT), sd)
colnames(sdRT_trim) = c("participant", "sdRT_trim")

RT_agg_trim = merge(meanRT_trim, sdRT_trim, by = "participant")

## merge aggregate info with long data
new_sem_z = merge(sem_z_trimmed, RT_agg_trim, by = "participant", all.x = T)

## person and grand-mean centered scores using original and aggregate
library(dplyr)
new_sem_z = new_sem_z %>% mutate(zRT_final = (RT - MeanRT_trim)/sdRT_trim)
                 
## checking: subject level means should be zero

sub_pic = group_by(new_sem_z, participant) %>%
  summarise_at(vars(zRT_final), mean)

```

## RTs per prime condition

## Partitioning the Data

```{r}
RTnotstudied = new_sem_z %>% filter(WordStatus == "Word was: New")
## now these "new" items could have been within/cross-trial primed
## although "within" is probably very few trials

## we look at zRTs of these "new" items

RTstudied = new_sem_z %>% filter(WordStatus == "Word was: Old")
## now these "old" items could have been within/cross-trial primed

## we look at zRTs of these "old" items
```

## Old (Studied) Items


### Within-Trial priming
```{r}
withintrial_rt = group_by(RTstudied, participant, WithinTrialPrimeCondition) %>% 
  summarise_at(vars(zRT_final), mean)


withintrial_rt_rmisc = Rmisc::summarySE(withintrial_rt,
                                     measurevar = "zRT_final",
                          groupvars = c( "WithinTrialPrimeCondition"))

withintrial_rt_rmisc = withintrial_rt_rmisc %>%
  filter(WithinTrialPrimeCondition != "firstword")

withintrial_rt_rmisc %>% 
  ggplot(aes(x = WithinTrialPrimeCondition, y = zRT_final, 
             group =WithinTrialPrimeCondition , fill = WithinTrialPrimeCondition))+
  #geom_point()+
geom_bar(stat = "identity", position = "dodge", width = 0.7)+
  geom_errorbar(aes(ymin=zRT_final - se, ymax=zRT_final + se), 
             width=.2, color = "gray26", 
             position = position_dodge(0.7))+
 theme_few()+
  scale_fill_gdocs()+
    xlab("") + ylab("z-RT") + 
  ggtitle("OLD items: within-trial priming")  +
   theme(axis.text = element_text(size = rel(1)),
          axis.title = element_text(face = "bold", size = rel(1)),
          legend.title = element_text(face = "bold", size = rel(1)),
         plot.title = element_text(hjust = .5),
         strip.text.x = element_text(face = "bold", size = rel(1.4)))

withintrial_rt_data = withintrial_rt %>% filter(!WithinTrialPrimeCondition %in% 
                                                  c("firstword", "Unprimed"))

withintrial_rt_aov= aov(data = withintrial_rt_data, 
                        zRT_final ~ WithinTrialPrimeCondition + 
                          Error(participant/WithinTrialPrimeCondition))
summary(withintrial_rt_aov)

withintrial_rt_ident = withintrial_rt_data %>% 
  filter(WithinTrialPrimeCondition == "Identical")

withintrial_rt_unrelated = withintrial_rt_data %>% 
  filter(WithinTrialPrimeCondition == "Unrelated")

withintrial_rt_rel = withintrial_rt_data %>% 
  filter(WithinTrialPrimeCondition == "Related")

t.test(withintrial_rt_ident$zRT_final, withintrial_rt_unrelated$zRT_final, paired = TRUE)

```
### Cross-Trial priming
```{r}
crosstrial_rt = group_by(RTstudied, participant, 
                         CrossTrialPrimeCondition) %>%
  summarise_at(vars(zRT_final), mean)

crosstrial_rmisc_rt = Rmisc::summarySE(crosstrial_rt,
                                     measurevar = "zRT_final",
                        groupvars = c( "CrossTrialPrimeCondition"))

crosstrial_rmisc_rt %>% 
  ggplot(aes(x = CrossTrialPrimeCondition, y = zRT_final, 
             group =CrossTrialPrimeCondition , fill = CrossTrialPrimeCondition))+
geom_bar(stat = "identity", position = "dodge", width = 0.7)+
  geom_errorbar(aes(ymin=zRT_final - se, ymax=zRT_final + se), 
             width=.2, color = "gray26", 
             position = position_dodge(0.7))+
 theme_few()+
  scale_fill_gdocs()+
    xlab("") + ylab("z-RT") + 
  ggtitle("OLD items: cross-trial priming")  +
   theme(axis.text = element_text(size = rel(1)),
          axis.title = element_text(face = "bold", size = rel(1)),
          legend.title = element_text(face = "bold", size = rel(1)),
         plot.title = element_text(hjust = .5),
         strip.text.x = element_text(face = "bold", size = rel(1.4)))
```

## New (Not Studied) Items


### Within-Trial priming
```{r}
withintrial_rt = group_by(RTnotstudied, participant, WithinTrialPrimeCondition) %>% 
  summarise_at(vars(zRT_final), mean)


withintrial_rt_rmisc = Rmisc::summarySE(withintrial_rt,
                                     measurevar = "zRT_final",
                          groupvars = c("WithinTrialPrimeCondition"))

withintrial_rt_rmisc %>% 
  ggplot(aes(x = WithinTrialPrimeCondition, y = zRT_final, 
             group =WithinTrialPrimeCondition , fill = WithinTrialPrimeCondition))+
  #geom_point()+
geom_bar(stat = "identity", position = "dodge", width = 0.7)+
  geom_errorbar(aes(ymin=zRT_final - se, ymax=zRT_final + se), 
             width=.2, color = "gray26", 
             position = position_dodge(0.7))+
 theme_few()+
  scale_fill_gdocs()+
    xlab("") + ylab("z-RT") + 
  ggtitle("NEW items: within-trial priming")  +
   theme(axis.text = element_text(size = rel(1)),
          axis.title = element_text(face = "bold", size = rel(1)),
          legend.title = element_text(face = "bold", size = rel(1)),
         plot.title = element_text(hjust = .5),
         strip.text.x = element_text(face = "bold", size = rel(1.4)))
```
### Cross-Trial priming
```{r}
crosstrial_rt = group_by(RTnotstudied, participant, CrossTrialPrimeCondition) %>%
  summarise_at(vars(zRT_final), mean)

crosstrial_rmisc_rt = Rmisc::summarySE(crosstrial_rt,
                                     measurevar = "zRT_final",
                        groupvars = c("CrossTrialPrimeCondition"))

crosstrial_rmisc_rt %>% 
  ggplot(aes(x = CrossTrialPrimeCondition, y = zRT_final, 
             group =CrossTrialPrimeCondition , fill = CrossTrialPrimeCondition))+
  #geom_point()+
geom_bar(stat = "identity", position = "dodge", width = 0.7)+
  geom_errorbar(aes(ymin=zRT_final - se, ymax=zRT_final + se), 
             width=.2, color = "gray26", 
             position = position_dodge(0.7))+
 theme_few()+
  scale_fill_gdocs()+
    xlab("") + ylab("z-RT") + 
  ggtitle("NEW items: cross-trial priming")  +
   theme(axis.text = element_text(size = rel(1)),
          axis.title = element_text(face = "bold", size = rel(1)),
          legend.title = element_text(face = "bold", size = rel(1)),
         plot.title = element_text(hjust = .5),
         strip.text.x = element_text(face = "bold", size = rel(1.4)))
```






